Revisiting the Truisms of Software Engineering:\\ Does Phase Delay Dramatically Increase Defect Fix Time?

### 2015-03-16 phaseDelay.tex
Updates from ShareLaTeX

\usepackage{tabularx}
\usepackage{subcaption}
\usepackage{picture}
\usepackage{wrapfig}
\usepackage{algorithm}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}} 
\newcommand{\lucas}[1]{\textcolor{red}{LUCAS: #1}} \newcommand{\tim}[1]{\textcolor{orange}{TIM: #1}}
\newcommand{\bill}[1]{\textcolor{blue}{BILL: #1}} 
\newcommand{\carter}[1]{\textcolor{cyan}{CARTER: #1}} 
\newcommand{\sei}[1]{\textcolor{RedViolet}{BILL/FORREST: #1}} 
\newcommand{\todo}[1]{\textcolor{Maroon}{TODO: #1}} 
%\newenvironment{changed}{\par}{\par}

@@

\usepackage{colortbl}
 \usepackage[font={small}]{caption, subfig}
\setlength{\abovecaptionskip}{1ex}
 \setlength{\belowcaptionskip}{1ex}

 \setlength{\floatsep}{1ex}
 \setlength{\textfloatsep}{1ex}
\usepackage[compact,small]{titlesec}
\DeclareMathSizes{7}{7}{7}{7} 
\pagenumbering{arabic}
\setlength{\columnsep}{7mm}

@@

\conferenceinfo{FSE}{'15 Bergamo, Italy}
\title{Revisiting the Truisms of Software Engineering:\\ Does Phase Delay Dramatically Increase  Repair Time?}
\numberofauthors{3}
\author{
\alignauthor
Tim Menzies, \\Carter Pape\\
       \affaddr{Dept. of Computer Science\\
       North Carolina State U, USA}\\
       tim.menzies@gmail.com,\\carterpape@gmail.com
\alignauthor
William Nichols,\\ Forrest Shull\\
\affaddr{Software Engineering Institute\\
        Carnegie Mellon U, USA}
wrn,fjshull@sei.cmu.edu
\alignauthor
Lucas \\Layman\\
       \affaddr{Fraunhofer CESE, \\
                College Park, USA}\\ 
       llayman@fc-md.umd.edu
}

\maketitle
\begin{abstract}
Does
repair time increase dramatically
the longer a defect persists in a system?
This widely held belief is used to justify 
major investments in  software development
 processes and methods.

This paper shows that this belief is maintained
quite strongly in the industrial software engineering
community (but, in the academic community, somewhat less so).
Yet based on a sample of 
171 software projects from around the world from 
and completed between 2006 and 2014, we argue that this belief is mostly 
incorrect. Specifically: the median fix time for issues
is just a few minutes and this does not tend to increase
the longer the issue remains in the system.

%This result raises the question: how did myths like phase delay survive so long  without  critical reassessment.  As a community, we need to reflect more on research practices and on how (and when) we reflect on the prominent hypotheses of our field.
\end{abstract}

@@

\vspace{1mm}
\noindent
{\bf Keywords:} phase delay, time to fix, agile, waterfall.

@@

Accordingly, this paper revisits
the {\em phase delay effect}
i.e.,  the rule of thumb stating that the time required to resolve an issue in a software project increases dramatically 
the longer it remains in the system. 
Phase delay has long been
a    motivator for major changes to
SE practices. Basili \& Boehm comment that since the 1980s,  phase delay:
\begin{quote}
...has been a major driver in focusing
industrial software practice on thorough
requirements analysis and design,
on early verification and validation, and
on up-front prototyping and simulation
to avoid costly downstream fixes~\cite{boehm01}.
\end{quote}
Basili and Boehm are not alone in this assessment of the significance of the phase delay effect. Our surveys of industrial practitioners (shown in \S\ref{survey}) reveal the phase-delay effect to be
the single most strongly held belief amongst commercial SE engineers.
Previously~\cite{me08a}, we have constructed return-on-investement models
for a common practice amongst U.S. government contractors: 
independent verification \& validation. IVV has been criticized as
ineffective since IVV engineers, many of whom are unfamiliar with the details of the project, are required to pass judgement on the code.
Hence,  the IVV team may find the same or fewer bugs
as the developers.
On the other hand, IVV has been defended
as follows:  when the IVV team finds bugs earlier than the developers then, due to the phase delay effect, this provides overall cost savings to the project. Note that the strength of this argument diminishes
if fixing later-phase issues is not significantly more expensive than fixing them earlier.

@@

In this paper, we investigate two hypotheses:

\bi
    \item $H_1$: The cost-to-fix of defects monotonically increases across phases.
    \item $H_2$: The cost-to-fix of defects monotonically increases as the phase delay increases.
\ei
The rest of the paper investigates these two hypotheses. First, we conduct a survey of practitioners and researchers to gauge their belief in the phase delay effect. Second, we  analyze 171 software  projects developed in the period 2006 to 2014.  To the best of our knowledge, this is the largest study on phase delay yet seen in the literature.

@@

\subsection{Why Study Phase Delay?}
\todo{\begin{itemize}
\item Given the Intro and Motivation, this section should focus on the influence of phase delay in software engineering milieux
\item Need to tone down the reliance on Beck. The important Beck point is that eliminating the cost-to-fix curve (resulting from phase delay) is one of the main goals of agile. His observations on new technologies are good too.
\item Need to cite other popular sources of this information, e.g., McConnell's "Ounce of Prevention" article~\cite{mcconnell01}.
\item Robert Glass's~\cite{glass02} book is very popular and contains an entry on this point
\item Google Scholar Citations for phase delay works Boehm88~\cite{boehm88} (673), Boehm \& Basili~\cite{boehm01} (761), 
\end{itemize}}

Due to its prominence in the field of software engineering,
this paper studies phase delay. As mentioned in the introduction, concerns about phase delay have
been used for decades to motivate major changes to software engineering practices. For example,
in 2000, Kent Beck used phase delay to motivate his call for ``extreme programming'' (a set of techniques
that evolved into contemporary agile and DevOps practices). Beck writes:
\begin{quote}
One of the universal assumptions of software engineering is that the cost of changing a program rises exponentially over time. I can remember sitting in a big linoleum-floored classroom as a college junior and seeing the professor draw on the board the curve found in \fig{curve1}.

The cost to fix a problem in a piece of software rises exponentially over time. A problem that might take a dollar to fix if you found it during requirements analysis might costs thousands to fix once the software is in production.

I resolved then and there that I would never let a problem get through to production. No sirree, I was going to catch problems as soon as possible. I would work out every possible problem in advance. I would review and crosscheck my code. No way was I going to cost my employer \$100,000.~\cite{beck00}
\end{quote}

@@

\begin{figure}[!t]
\scriptsize
\begin{center}
\begin{tabular}{l@{~}|l@{~}|r@{~}r@{~}|r@{~}l}
          &        & \multicolumn{2}{c|}{Time to}  &\\
          &        & \multicolumn{2}{c|}{resolve} &\\
Injection & Removal& initial & now & \multicolumn{2}{l}{Scale up w.r.t initial} \\\hline
Phase $i$&Phase $i$ & 1 & 1 & 1 & **\\
Phase $i$ &Phase $i+1$ & 1 & 2& 2 & **** \\
Phase $i$ &Phase $i+2$ & 1 & 4& 4 &******** \\
Phase $i$ &Phase $i+3$ & 1 & 8 & 8 & ****************  
  \end{tabular}\end{center}
 \caption{Projects data consistent with
 the phase delay effect of \fig{curve1}.}\label{fig:curve1a}
\end{figure}

\begin{figure}[!t]
\scriptsize
\begin{center}
\begin{tabular}{l@{~}|l@{~}|r@{~}r@{~}|r@{~}l}
          &        & \multicolumn{2}{c|}{Time to}  &\\
          &        & \multicolumn{2}{c|}{resolve} &\\
Injection & Removal& initial & now & \multicolumn{2}{l}{Scale up w.r.t initial} \\\hline
Phase $i$&Phase $i$ & 1 & 1 & 1 & ** \\
Phase $i$ &Phase $i+1$ & 1 & 2& 2 & **** \\
Phase $i$ &Phase $i+2$ & 1 & 2& 2 & **** \\
Phase $i$ &Phase $i+3$ & 1 & 2 & 2 & **** \\
 Phase $i$  &Phase $i+4$ &1 & 2& 2 &**** 
  \end{tabular}\end{center}
 \caption{Projects data consistent with
 the phase delay effect of \fig{curve2}.}\label{fig:curve2a}
\end{figure}

Beck speculated that there might be another kind of project, shown in \fig{curve2},
that does not exhibit phase delay. In this next quote, Beck clearly states
the importance of avoiding phase delay (in fact, for Beck, it is {\em the} major
motivation for his work):
\begin{quote}
This (\fig{curve2}) is one of the premises of (agile). It is the technical premise of (agile). If the cost of change rose slowly over time, you would act completely differently from how you do under the assumption that costs rise exponentially. You would make big decisions as late in the process as possible, to defer the cost of making the decisions and to have the greatest possible chance that they would be right. You would only implement what you had to, in hopes that the needs you anticipate for tomorrow wouldn't come true. You would introduce elements to the design only as they simplified existing code or made writing the next bit of code simpler.~\cite{beck00}
\end{quote}
Beck makes one other comment that is relevant to this paper: 
\begin{quote}
The software development community has spent enormous resources in recent decades trying to reduce the cost of change-better languages, better database technology, better programming practices, better environments and tools, new notations.

@@

\todo{
\bi
    \item What is the cost-to-fix curve in TSP data? Check against H1 and compare to past literature.
    \item What is the effect of phase delay? Check against H2.
    \item \fig{cost-to-fix-tsp} is great. really important. its from  different decades, right?  so can we add text saying that the general pattern is that the (approx) 15 years between each curve tends to a flattening. not, perhaps, due to process but to better build environs.
\ei
}

@@

The distribution of defects found and fixed per phase in our data is shown in Figure~\ref{fig:fix-phase-dist}. A high percentage of defects were found and fixed in the early phases, i.e., requirements, high level design, and design reviews and inspections. \todo{How does this compare evidence from other sources?}

The median time to fix a defect in each phase is shown in \fig{fix-time-per-phase}. The median cost to fix a defect is low in the early phases, and then increases slightly during design and coding before peaking during the unit test phase. The median fix time then drops during the testing phases. Note that while the median time remains relatively consistent, the 75th percentile does increase during the testing phases.

@@

\begin{figure}[!ht]
\begin{center}
\begin{tabular}{r|rrr|l}
  Sample&\multicolumn{3}{c|}{Percentiles}\\ 
size & 25th & 50th & 75th & Phase removed \\
\hline
120 & 3 & 3 & 18 & ReqtsInspect \\
126 & 3 & 3 & 20 & ReqtsReview \\
70 & 2 & 2 & 10 & HLDInspect \\
 55 &    2 &    1&    7&HLDReview \\

278&    7&    7&   33&DesignInspect \\
233&    6&    6&   25&DesignReview \\

315&    8&    7&   34&CodeInspect \\ 
282&    7&    7&   32&CodeReview\\

425&   11&   10&   54&UnitTest \\ 
226&    7&    6&   46&IntTest \\ 
194&    7&    6&   52&SysTest\\ 
 88&    4&    3&   38&AcceptTest\\
\end{tabular}
\end{center}
\caption{Median fix time per phase}
\label{fig:fix-time-per-phase}
\end{figure}

For illustrative purposes, the relative cost to fix errors per lifecycle phase in the TSP project data is overlayed with Boehm's cost curve~\cite{Boehm81} as well as the CCPDS-R case study~\cite{Royce98} in \fig{cost-to-fix-tsp}. It is clear to see that the trend is relatively flat, closer to the pattern exhibited by the Royce case study (which also focused on early defect removal) than the exponential growth curve.

@@

We reject $H_1$.

\subsection{Phase delay analysis}
\todo{
\bi
    \item What can we say about \fig{raw}
\ei
}

Our second hypothesis was that defects that exhibit higher phase delay (i.e. that persist in the software over more lifecycle phases) tend to require more effort to fix than those with less phase delay. To investigate this, we calculate phase delay as simply the number of project phases (the set used in our data set is shown in \fig{waterfall}) during which the defect existed in the system, i.e. the ordinal rank of removal phase minus the ordinal rank of injection phase. Sub-phases (such as review and inspection) are not included in this calculation.

\fig{correlation} and \fig{phase_delay_medians} show that most defects have a relatively small phase delay in our project sample (phase delay is often <= 2). As intended, the Review and Inspection Phases for Requirements, Design, and Code mostly find defects originating in the same phase. In contrast, the Testing phases mostly find defects originating in the Code phase.

@@

We reject $H_2$.

%In the TSP data, there is a weak correlation (Kendall's $\tau = 0.19, p < 0.01, 95\% CI$) between a defect's cost-to-fix (minutes of effort) and the defects phase delay (phase removed - phase injected).

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=3.3in]{corr.pdf}
\end{center}
\caption{Correlation of phase delay and fix time (plot shows a 1\% sampling of data).}
\label{fig:correlation}
\end{figure}

\begin{figure}[!ht]
\centering
\begin{tabular}{rrr|l}
    \multicolumn{3}{c|}{Percentiles}\\ 
  25th & 50th & 75th & Phase removed \\ 
  \hline
    2 & 2 & 2 & ReqtsInspect \\ 
    1 & 1 & 1 & ReqtsReview \\ 
    2 & 2 & 2 & HLDInspect \\ 
    1 & 1 & 1 & HLDReview \\ 
    2 & 2 & 2 & DesignInspect \\ 
    1 & 1 & 1 & DesignReview \\ 
    2 & 2 & 2 & CodeInspect \\ 
    1 & 1 & 1 & CodeReview \\ 
    3 & 3 & 6 & UnitTest \\ 
    4 & 4 & 4 & IntTest \\ 
    5 & 5 & 5 & SysTest \\ 
    6 & 6 & 6 & AcceptTest \\ 
  \end{tabular}
\caption{Median phase delay of defects removed per phase} 
\label{fig:phase_delay_medians}
\end{figure}

% \begin{figure}[ht]
% \centering
% \begin{tabular}{lll}
%   Phase & $tau$ & $p$ \\ 
%   \hline
%   Design & 0.14 & 0.08 \\ 
%   DesignReview & 0.06 & $<$0.01 \\ 
%   DesignInspect & 0.05 & $<$0.01 \\ 
%   Code & 0.27 & $<$0.01 \\ 
%   CodeReview & 0.12 & $<$0.01 \\ 
%   CodeInspect & 0.09 & $<$0.01 \\ 
%   Test & 0.13 & $<$0.01 \\ 
%   IntTest & 0.1 & $<$0.01 \\ 
%   SysTest & 0.25 & $<$0.01 \\ 
%   AcceptTest & 0.06 & 0.29 \\ 
%   \end{tabular}
% \caption{Correlation of cost-to-fix and phase delay for defects found (Kendall's $\tau$)} 
% \label{fig:phase_corr}
% \end{figure}

% \begin{figure*}[!ht]
% \centering
% \includegraphics[width=1.6in]{corr-Code.png}
% \includegraphics[width=1.6in]{corr-CodeReview.png}
% \includegraphics[width=1.6in]{corr-CodeInspect.png}
% \vspace{0.01cm}

% \includegraphics[width=1.6in]{corr-Design.png}
% \includegraphics[width=1.6in]{corr-DesignReview.png}
% \includegraphics[width=1.6in]{corr-DesignInspect.png}
% \vspace{0.01cm}

% \includegraphics[width=1.6in]{corr-Test.png}
% \includegraphics[width=1.6in]{corr-IntTest.png}
% \includegraphics[width=1.6in]{corr-SysTest.png}
% \includegraphics[width=1.6in]{corr-AcceptTest.png}

% \caption{Correlation of cost-to-fix (rank) and phase delay for defects found by phase for phases where \# defects >= 30 and \# unique phase delay values >= 5. x = phase delay, y = cost-to-fix rank. Plots use random samples from each phase for readability.}
% \label{fig:phase_fixtime_plot}
% \end{figure*}

>>>>

\usepackage{tabularx}
\usepackage{subcaption}
\usepackage{picture}
\usepackage{wrapfig}
\usepackage{algorithm}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}} 
\newcommand{\lucas}[1]{\textcolor{red}{LUCAS: #1}} 
\newcommand{\tim}[1]{\textcolor{Red}{TIM: #1}}
\newcommand{\bill}[1]{\textcolor{blue}{BILL: #1}} 
\newcommand{\carter}[1]{\textcolor{cyan}{CARTER: #1}} 
\newcommand{\sei}[1]{\textcolor{RedViolet}{BILL/FORREST: #1}} 
\newcommand{\todo}[1]{\textcolor{Maroon}{TODO: #1}} 
%\newenvironment{changed}{\par}{\par}

@@

\usepackage{colortbl}
 \usepackage[font={small}]{caption, subfig}
\setlength{\abovecaptionskip}{1ex}
 \setlength{\belowcaptionskip}{1ex}
 \setlength{\floatsep}{1ex}
 \setlength{\textfloatsep}{1ex}
\usepackage[compact,small]{titlesec}
\DeclareMathSizes{7}{7}{7}{7} 
\pagenumbering{arabic}
\setlength{\columnsep}{7mm}

@@

\conferenceinfo{FSE}{'15 Bergamo, Italy}
\title{Revisiting the Truisms of Software Engineering:\\ Does Phase Delay Dramatically Increase Defect Fix Time?}
\numberofauthors{5}
\author{
\alignauthor 
Tim Menzies \\
       \affaddr{Dept. of Computer Science}\\
       \affaddr{North Carolina State U, USA}\\
       \email{tim.menzies@gmail.com}
\alignauthor 
William Nichols \\
        \affaddr{Software Engineering Institute}\\
        \affaddr{Carnegie Mellon U, USA}
        \email{wrn@sei.cmu.edu} 
\alignauthor 
Lucas Layman \\
       \affaddr{Fraunhofer CESE} \\
       \affaddr{College Park, USA}\\ 
       \email{llayman@fc-md.umd.edu}
\and
\alignauthor 
Forrest Shull \\
        \affaddr{Software Engineering Institute}\\
        \affaddr{Carnegie Mellon U, USA}
        \email{fjshull@sei.cmu.edu} 
\alignauthor 
Carter Pape \\
       \affaddr{Dept. of Computer Science}\\
       \affaddr{North Carolina State U, USA}\\
       \email{carterpape@gmail.com}
} 
\maketitle
\begin{abstract}
Does
repair time increase dramatically
the longer a defect persists in a system?
This widely held belief is used to justify 
major investments in  software development
 processes and methods.

This paper shows that this belief is maintained
quite strongly in the industrial software engineering
community (but, in the academic community, a little less so).
Yet based on a sample of 
171 software projects from around the world from 
and completed between 2006 and 2014, we argue that this belief is mostly 
incorrect. Specifically: the median fix time spent in phases
devoted to resolving issues does {\em not} increase
the longer the issue remains in the system.

This result raises the question: how do widely-held beliefs like phase delay survive so long  without  critical reassessment?  As a community, we need to reflect more on research practices and on how (and when) we reflect on the prominent hypotheses of our field.
\end{abstract}

@@

\vspace{1mm}
\noindent
{\bf Keywords:} software processes, economics, phase delay, time to fix.

@@

Accordingly, this paper revisits
the {\em phase delay effect}
i.e.,  the rule of thumb stating that the time required to resolve an issue in a software project increases dramatically 
the longer it remains in the system. 
Phase delay has long been
a    motivator for major changes to
SE practices. Basili \& Boehm comment that since the 1980s,  phase delay:
\begin{quote}
...has been a major driver in focusing
industrial software practice on thorough
requirements analysis and design,
on early verification and validation, and
on up-front prototyping and simulation
to avoid costly downstream fixes~\cite{boehm01}.
\end{quote}
Basili and Boehm are not alone in this assessment of the significance of the phase delay effect. Our surveys of industrial practitioners (shown in \S\ref{survey}) reveal the phase delay effect to be
one of the most strongly-held beliefs amongst commercial software engineers.
Previously~\cite{me08a}, we have constructed return-on-investment models
for a common practice amongst U.S. government contractors: 
independent verification \& validation. IVV has been criticized as
ineffective since IVV engineers, many of whom are unfamiliar with the details of the project, are required to pass judgement on the code.
Hence,  the IVV team may find the same or fewer bugs
as the developers.
On the other hand, IVV has been defended
as follows:  when the IVV team finds bugs earlier than the developers then, due to the phase delay effect, this provides overall cost savings to the project. Note that the strength of this argument diminishes
if fixing later-phase issues is not significantly more expensive than fixing them earlier.

@@

In this paper, we investigate several hypotheses:

\bi
    \item $H_1$: The cost-to-fix of defects monotonically increases across phases.
    \item $H_2$: The cost-to-fix of defects monotonically increases as the phase delay increases.
\ei
As we explored the above, we found a related hypothesis was believed by both professional software engineers and FSE conference reviewers:
\bi
 \item $H_3$: Requirements errors are the most expensive to fix.
\ei

The rest of the paper investigates these  hypotheses. First, we conduct a survey of practitioners and researchers to gauge their belief in the phase delay effect. Second, we  analyze 171 software  projects developed in the period 2006 to 2014.  To the best of our knowledge, this is the largest study on phase delay yet seen in the literature.

@@

\subsection{Why Study Phase Delay?}
\label{sec:why-study}
%\todo{\begin{itemize}
%\item Given the Intro and Motivation, this section should focus on the influence of phase delay in software engineering milieux
%\item Need to tone down the reliance on Beck. The important Beck point is that eliminating the cost-to-fix curve (resulting from phase delay) is one of the main goals of agile. His observations on new technologies are good too.
%\item Need to cite other popular sources of this information, e.g., McConnell's "Ounce of Prevention" article~\cite{mcconnell01}.
%\item Robert Glass's~\cite{glass02} book is very popular and contains an entry on this point
%\item Google Scholar Citations for phase delay works Boehm88~\cite{boehm88} (673), Boehm \& Basili~\cite{boehm01} (761), 
%\end{itemize}}

This paper examines evidence for the phase delay effect due to its prominence throughout the field of software engineering. Many prominent authors have commented on its perceived usefulness as a rule of thumb for software engineers. For example, in 2001, Steve McConnell mentioned it as a "common observation" in the field and proceeded to knock down arguments that it no longer held for contemporaneous software projects, by relying on recently published data and experience. McConnell also summarized the intuitive argument for why it should be so: ``A small mistake in upstream work can affect large amounts of downstream work. A change to a single sentence in a requirements specification can imply changes in hundreds of lines of code spread across numerous classes or modules, dozens of test cases, and numerous pages of end-user documentation''~\cite{mcconnell01} Glass echoes this sentiment, stating the the fact ``requirements errors are the most expensive to fix when found during production but the cheapest to fix early in development'' is ``really just common sense''~\cite{glass02}.

Also in 2001, the phase delay effect was listed as number one by Boehm and Basili in their "Top 10 list" of "objective and quantitative data, relationships,
and predictive models that help
software developers avoid predictable pitfalls
and improve their ability to predict
and control efficient software projects." \cite{boehm01}

More important than its use as a rule of thumb, the phase delay effect has 
been used for decades to motivate major changes to software engineering practices (as mentioned in Section 1). For example,
in 2000, Kent Beck used phase delay to motivate his call for ``extreme programming'' (a set of techniques
that evolved into contemporary agile and DevOps practices). Beck writes:
\begin{quote}
One of the universal assumptions of software engineering is that the cost of changing a program rises exponentially over time. I can remember sitting in a big linoleum-floored classroom as a college junior and seeing the professor draw on the board the curve found in \fig{curve1}... 
%
%The cost to fix a problem in a piece of software rises exponentially over time. 
A problem that might take a dollar to fix if you found it during requirements analysis might costs thousands to fix once the software is in production.
%
%I resolved then and there that I would never let a problem get through to production. No sirree, I was going to catch problems as soon as possible. I would work out every possible problem in advance. I would review and crosscheck my code. No way was I going to cost my employer \$100,000.~
\cite{beck00}
\end{quote}

@@

% \begin{figure}[!t]
% \scriptsize
% \begin{center}
% \begin{tabular}{l@{~}|l@{~}|r@{~}r@{~}|r@{~}l}
%           &        & \multicolumn{2}{c|}{Time to}  &\\
%           &        & \multicolumn{2}{c|}{resolve} &\\
% Injection & Removal& initial & now & \multicolumn{2}{l}{Scale up w.r.t initial} \\\hline
% Phase $i$&Phase $i$ & 1 & 1 & 1 & **\\
% Phase $i$ &Phase $i+1$ & 1 & 2& 2 & **** \\
% Phase $i$ &Phase $i+2$ & 1 & 4& 4 &******** \\
% Phase $i$ &Phase $i+3$ & 1 & 8 & 8 & ****************  
%   \end{tabular}\end{center}
%  \caption{Projects data consistent with
%  the phase delay effect of \fig{curve1}.}\label{fig:curve1a}
% \end{figure}

% \begin{figure}[!t]
% \scriptsize
% \begin{center}
% \begin{tabular}{l@{~}|l@{~}|r@{~}r@{~}|r@{~}l}
%           &        & \multicolumn{2}{c|}{Time to}  &\\
%           &        & \multicolumn{2}{c|}{resolve} &\\
% Injection & Removal& initial & now & \multicolumn{2}{l}{Scale up w.r.t initial} \\\hline
% Phase $i$&Phase $i$ & 1 & 1 & 1 & ** \\
% Phase $i$ &Phase $i+1$ & 1 & 2& 2 & **** \\
% Phase $i$ &Phase $i+2$ & 1 & 2& 2 & **** \\
% Phase $i$ &Phase $i+3$ & 1 & 2 & 2 & **** \\
%  Phase $i$  &Phase $i+4$ &1 & 2& 2 &**** 
%   \end{tabular}\end{center}
%  \caption{Projects data consistent with
%  the phase delay effect of \fig{curve2}.}\label{fig:curve2a}
% \end{figure}

Beck speculated that there might be another kind of project, shown in \fig{curve2},
that does not exhibit phase delay. In fact, for Beck, the possibility that phase delay does not imply exponential increases in the resources required to make changes and fix defects is {\em the} major
motivation for his work, because it means that decisions can be made as late in the process as possible, so as to increase the chances that these decisions would be right. Beck found more than adequate motivation for this new assumption:
%\begin{quote}
%This (\fig{curve2}) is one of the premises of (agile). It is the technical premise of (agile). If the cost of change rose slowly over time, you would act completely differently from how you do under the assumption that costs rise exponentially. You would make big decisions as late in the process as possible, to defer the cost of making the decisions and to have the greatest possible chance that they would be right. You would only implement what you had to, in hopes that the needs you anticipate for tomorrow wouldn't come true. You would introduce elements to the design only as they simplified existing code or made writing the next bit of code simpler.~\cite{beck00}
%\end{quote}
%Beck makes one other comment that is relevant to this paper: 
\begin{quote}
The software development community has spent enormous resources in recent decades trying to reduce the cost of change-better languages, better database technology, better programming practices, better environments and tools, new notations.

@@

%\todo{
%\bi
%    \item What is the cost-to-fix curve in TSP data? Check against H1 and compare to past literature.
%    \item What is the effect of phase delay? Check against H2.
%    \item \fig{cost-to-fix-tsp} is great. really important. its from  different decades, right?  so can we add text saying that the general pattern is that the (approx) 15 years between each curve tends to a flattening. not, perhaps, due to process but to better build environs.
%\ei
%}

@@

The distribution of defects found and fixed per phase in our data is shown in Figure~\ref{fig:fix-phase-dist}. A high percentage of defects were found and fixed in the early phases, i.e., requirements, high level design, and design reviews and inspections. (This distribution is similar to that observed for other projects that emphasized investment in software engineering quality assurance practices, for example the NASA robotics projects which had 58\% of defects removed before implementation began. \cite{me08a} Similar results are presented by Jones and Bonsignour, who show 52\% of pretest defects removed before entering implementation. \cite{jones12})

\todo{This paragraph is wrong now.} The median time to fix a defect in each phase is shown in \fig{fix-time-per-phase}. The median cost to fix a defect is low in the early phases, and then increases slightly during design and coding before peaking during the unit test phase. The median fix time then drops during the testing phases. Note that while the median time remains relatively consistent, the 75th percentile does increase during the testing phases.

@@

\begin{figure*}[!ht]

 \renewcommand{\baselinestretch}{0.7}
 \scriptsize
\begin{center}
\begin{tabular}{r|rr|rl}
Phase & samples & total time&\multicolumn{2}{l}{Total time w.r.t. first phase}\\\hline
~\\
Planning&229&12625&1.00&*\\
~\\
Reqts&590&24182&1.91&**\\
ReqtsInspect&2533&92723&7.34&********\\
ReqtsReview&1762&67471&5.34&******\\
~\\
HLD&78&6856&0.54&*\\
HLDReview&472&12821&1.01&**\\
HLDInspect&1366&35291&2.79&***\\
~\\
Design&620&21573&1.70&**\\
DesignInspect&8373&229937&18.21&*******************\\
DesignReview&3043&92823&7.35&********\\
~\\
Code&3029&220058&17.43&******************\\
CodeReview&4774&163091&12.91&*************\\
CodeInspect&11406&334442&26.49&***************************\\
~\\
UnitTest&9177&352245&2.790&****************************\\
IntTest&2761&160559&12.71&*************\\
SysTest&2059&105796&8.37&*********\\
AcceptTest&1045&42717&0.338&****\\
\end{tabular}
\end{center}
\caption{Times per phase (minutes).}
\label{fig:fix-time-per-phase}
\end{figure*}

For illustrative purposes, the relative cost to fix errors per lifecycle phase in the TSP project data is overlayed with Boehm's cost curve~\cite{Boehm81} as well as the CCPDS-R case study~\cite{Royce98} in \fig{cost-to-fix-tsp}. Cost-to-fix for TSP is total time spent in each review, inspection, or test phase (reported in \fig{fix-time-per-phase}) / \# of defects found in phase (reported in \fig{raw}). It is clear to see that the trend is relatively flat, closer to the pattern exhibited by the Royce case study (which also focused on early defect removal) than the exponential growth curve.

@@

Given the above evidence, we reject $H_1$: The cost-to-fix of defects monotonically increases across phases.

\subsection{Phase delay analysis}

Our second hypothesis was that defects that exhibit higher phase delay (i.e. that persist in the software over more lifecycle phases) tend to require more effort to fix than those with less phase delay. To investigate this, we calculate phase delay as simply the number of project phases (the set used in our data set is shown in \fig{waterfall}) during which the defect existed in the system, i.e. the ordinal rank of removal phase minus the ordinal rank of injection phase. Sub-phases (such as review and inspection) are included in this calculation.

@@

Given the above evidence, we reject $H_2$: the cost-to-fix of defects monotonically increases as the phase delay increases.

\begin{figure*}[!t] 
 \renewcommand{\baselinestretch}{0.7}
 \scriptsize
\begin{center}
\begin{tabular}{r|rrr|ll|rl}
  Sample&\multicolumn{3}{c|}{Percentiles}\\ 
size & 25th & 50th & 75th & Phase injected & Phase removed & \multicolumn{2}{l}{Scale up w.r.t. to first phase}\\\hline
\\
 48&   14&   30&   51&Before&DesignInspect &1.00 & **********  \\
 94&   22&   45&   80&Before&CodeInspect &1.50 & ***************  \\
 53&   13&   28&   49&Before&CodeReview &0.93 & **********  \\
153&   38&   77&  130&Before&UnitTest &2.56 & **************************  \\
127&   30&   65&  128&Before&IntTest &2.16 & **********************  \\
 92&   24&   56&   89&Before&SysTest &1.86 & *******************  \\

\\
 56&   14&   35&   57&Planning&ReqtsReview &1.00 & **********  \\
 46&   11&   24&   39&Planning&DesignInspect &0.68 & *******  \\
 32&   18&   33&   84&Planning&UnitTest &0.94 & **********  \\

\\
189&   46&   93&  155&Reqts&ReqtsInspect &1.00 & **********  \\
184&   45&   92&  155&Reqts&ReqtsReview &0.98 & **********  \\
102&   24&   49&   83&Reqts&DesignInspect &0.52 & ******  \\
 61&   14&   33&   55&Reqts&DesignReview &0.35 & ****  \\
 59&   18&   38&   60&Reqts&CodeInspect &0.40 & ****  \\
 91&   22&   49&  104&Reqts&UnitTest &0.52 & ******  \\
100&   33&   67&  120&Reqts&IntTest &0.72 & ********  \\
 78&   28&  122&  237&Reqts&SysTest &1.31 & **************  \\
\\
207&   50&  101&  156&Design&DesignInspect &1.00 & **********  \\
210&   51&  103&  166&Design&DesignReview &1.01 & ***********  \\
163&   39&   79&  135&Design&CodeInspect &0.78 & ********  \\
134&   32&   66&  116&Design&CodeReview &0.65 & *******  \\
238&   58&  119&  183&Design&UnitTest &1.17 & ************  \\
172&   42&   88&  158&Design&IntTest &0.87 & *********  \\
163&   40&   80&  161&Design&SysTest &0.79 & ********  \\
 80&   21&   45&   73&Design&AcceptTest &0.44 & *****  \\

\\
130&   31&   65&  122&Code&AcceptTest &1.00 & **********  \\
228&   56&  113&  176&Code&CodeInspect &1.73 & ******************  \\
235&   57&  115&  177&Code&CodeReview &1.76 & ******************  \\
243&   59&  125&  206&Code&IntTest &1.92 & ********************  \\
191&   46&   95&  173&Code&SysTest &1.46 & ***************  \\
269&   66&  133&  206&Code&UnitTest &2.04 & *********************  \\
 \end{tabular}
\end{center}
\caption{Distribution of fix times seen in SEI TSP data.}
\label{fig:raw}
\end{figure*}
